Evaluating Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe using radiocarbon data

### 2019-03-17 review/answer_to_review.md

## Comments to the Author:

> A lot of work has been already done in quantifying changes in funerary practices during the Bronze Age in central and south-western Europe through the analysis of radiocarbon-dated archaeological contexts. I recommend citing these works in the paper, since the adopted approach is similar.

> -	Capuzzo G, Barceló JA. 2015. Cultural changes in the 2nd millennium BC: a Bayesian examination of radiocarbon evidence from Switzerland and Catalonia. World Archaeology 47 (4) Special Issue: Bayesian approaches to Prehistoric Chronologies:622-641.
> -	Capuzzo G, López Cachero FJ. 2017. De la inhumación a la cremación en el nordeste peninsular: cronología y sociedad. In: Barceló JA, Bogdanovic I, Morell B, editors. Iber-Crono 2016. Cronometrías Para la Historia de la Península Ibérica (Chronometry for the History of the Iberian Peninsula), Barcelona (Spain), 17-19 October 2016:192-208.
> -	Capuzzo G, Barceló JA (in press). Cremation burials in Central and Western Europe: quantifying an adoption of innovation in the 2nd millennium BC. In: Kneisel J, Nakoinz O, Barcelo J, editors. Turning Points and Change in Bronze Age Europe (2400 – 800 BC). Modes of change – inhumation versus cremation in Bronze Age burial rites. Proceedings of the International Workshop "Socio-Environmental Dynamics over the Last 12,000 Years: The Creation of Landscapes IV (24th-27th March 2015) in Kiel. Universitätsforschungen zur Prähistorischen Archäologie 
> -	Capuzzo G. 2014. Space-temporal Analysis of Radiocarbon Evidence and Associated Archaeological Record: from Danube to Ebro Rivers and from Bronze to Iron Ages. Ph.D. thesis, Autonomous University of Barcelona.

I was not aware of this work and added a paragraph in the introduction to reference it. It is indeed relevant.

@@

> In the paper, it’s very high the amount of data, for which the archaeological information regarding the type of burial and funerary rite is lacking. At table 1, page 4, we can clearly detect that for 60% of the dates associated to cremation burials we have no information regarding the type of funerary structure. For 37% of the inhumations, no information about the funerary structure is available. Moreover, for 32% of the dataset (559 radiocarbon dates out of 1701) we do not have any kind of information regarding neither the funerary rite nor the burial structure. This lack of data obliges the author to consider carefully the results reached in the paper. It’s relevant to highlight all the biases of the prosed research and how the conclusion can be misleading due to the quality of available data. In particular, when the statistical analyses carried out on data aim to explore in the details the cultural distance among the different regions.

This is absolutely true and I tried to emphasise the limitations of the data in the discussion. The challenges are not confined on the amount of data, but also concern its meaning and origin. The supplementary remarks are extensively long to cover these problems. Is it necessary to mention the limitations at other critical points of the paper? I'm absolutely open to do that.

@@

> Page 5, Column 1, Lines 24-60. I have several issues with this model. The idea here is to emulate the observed data so we can build expectations that can be directly compared against empirical patterns. But I do find several issues: 1) How is Ng=100 justified? This seems a somewhat arbitrary number for a parameter that is likely going to play a significant role in the amount of drift in the system? 2) I understand each time-step is 20 years. Why? I assume this is related to generation length but this ought to be discussed; 3) Is the model supposed to reach an equilibrium state? If this is the case I don’t think 70 time-steps and Ng=100 is sufficient for this. If the answer is no, the initial condition of the system shouldn’t be random and perhaps be informed from the empirical data.

> Figure 3. These are very nice plots - I would add the sample sizes for each region though.

> Figure 6. I find these plots a bit messy and with too much information. I also don’t see much the point of testing for significance the results of each simulation run - they are effectively samples from the same generative process so I think testing them does not make much sense?

Figure 6 is much simpler now without simulation results. It only contains a time series of distance correlation.

### 2019-03-24 review/answer_to_review.md

> Firstly, my impression is that the author is trying to squeeze perhaps a bit too much in a single paper. I felt this was particularly the case of the computer simulation model which is almost thrown in in the mix, without much discussion on model parameters (see below). I would be tempted to suggest removing this entirely and include this into another paper.

After careful deliberation, I have decided to follow this advice. The simulation (1) does not add information that would be absolutely essential for the core topics, (2) reduces the overall comprehensibility due to the terminology that has to be introduced to describe it and (3) cannot be adequately presented and discussed here without leaving the initial scope of the paper. I will prepare another paper in the future that will present this simulation and other simulation approaches.

> Secondly, I feel like the empirical analyses have too much preprocessing. The raw data provides spatial and temporal coordinates of individual burials, yet these are combined spatially and chronologically first and analysed as aggregates. As a consequence: 1) there is a substantial loss of spatial information; 2) sampling error is effectively disregarded (i.e. mantel tests are based on proportion values that are calculated for regions with very different samples sizes) and strictly speaking the samples are also non-independent (i.e. a large site with multiple burials can strongly bias the proportions of a particular region); and 3) there is an increased risk of ecological fallacy/Simpson’s paradox. I take this aggregation provides a direct mapping to existing literature, but I am not convinced that this is worth sacrificing so much.

The decision to create discrete spatial area classes is necessary to establish subsets of dates that can be used for time series calculation. I'm not aware of a methodologival approach where time series could be constructed without some sort of aggregation. Unfortunately the size of the areas has to be very big because of the currently overall small amount of data. I expect this to change in the future and the analysis can be redone then with more data and much smaller areas to understand the definitely much more complex processes within the macro regions established here.

So: The criticism listed here is valid and I added a paragraph in the *Data quality* section to mention it. I refer to this section multiple times in the article. Nevertheless I don't see a practical possibility

@@

> Thirdly, the author implies that the scripts used for his analysis are supplied but there are no links on the article to any repository and I did not find any electronic supplementary material. As the author is well aware, we need to do our best to make our work fully reproducible.

This is the result of a misunderstanding in the communication with the special issue editors. I added the link to the relevant repository now.

@@

## Minor points:

> Page 1, Column 2, Lines 52-43: “Do time series ... spatial distance”. I think this question should be phrased better. Eg. What  does “meaningful” mean in this context?

Rephrased to a more direct question: "How does the cultural distance between regions that results from changes in burial traditions develop, and how is this cultural distance related to spatial distance?".

### 2019-03-24 article/01_Introduction.Rmd
rephrased second main question

If radiocarbon data allows to reconstruct the diffusion of burial traditions, then the quantitative comparison of macro regional trends could also pave the way for an analysis of interaction intensity on a long-term continental scale. The second main question: *Do time series constructed from radiocarbon data enable meaningful observations concerning cultural and spatial distance?*

>>>>

If radiocarbon data allows to reconstruct the diffusion of burial traditions, then the quantitative comparison of macro regional trends could also pave the way for an analysis of interaction intensity on a long-term continental scale. The second main question: *How does the cultural distance between regions that results from changes in burial traditions develop, and how is this cultural distance related to spatial distance?*

### 2019-03-24 article/04_Discussion.Rmd
rephrased second main question

The second important question posed for this paper was, *if these time series constructed from radiocarbon data* (Figure \ref{development_burial_type} and \ref{development_burial_construction}) *allow further analysis concerning cultural and spatial distance*? That can also be affirmed. While the idea behind the first question was rather technical -- reproducing already available information in a quantitative way -- the second one had a larger scope. The quantitative representation of an archaeological narrative paths the way for a diverse set of methods to be applied and the analysis undertaken here is merely the beginning.

>>>>

The second important question posed for this paper was, *how cultural distance between regions that results from changes in burial traditions develops, and how it is related to spatial distance?* While the idea behind the first question was rather technical -- reproducing already available information in a quantitative way -- the second one had a larger scope. The quantitative representation of an archaeological narrative paths the way for a diverse set of methods to be applied and the analysis undertaken here is merely the beginning.
